---
title: "Infection during pregnancy and pregancy outcomes and anthropometrics, UCL 2023-24"
author: "Mathilde Le Vu, Jonathan Wells, Mario Cortina-Borja"
date: "2023-12-05"
html_document: 
output:
  rmdformats::robobook:
    code_folding: show
    self_contained: true
    highlight: pygments
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


# Anthropometric parameters
## Birthweight
#### M_BW: Livebirths
#### Adjusting for gestational age
```{r}
MBW.Z.GA2<- gam(Birthweight_Z_GA_LMP ~ s(birthdate_num)+ s(day366, bs="cc") + s(height)+  s(age_mother)+  flu_or_syph+parity_cat2+sex+morphology +civil_status_cat2+Lausanne, data=laus4_livebirths, subset=(GA_weeks >=22))
  summary(MBW.Z.GA2)
  BIC(MBW.Z.GA2)
  gam.check(MBW.Z.GA2)
  concurvity(MBW.Z.GA2)
```

##### Summary table
```{r}
BW_GAM_mod_adj_GA <- round(head(estimates_GAM_cont(MBW.Z.GA2),11), 2)
BW_GAM_mod_adj_GA <- BW_GAM_mod_adj_GA %>%
mutate(pvalue=as.numeric(pvalue)) %>%
mutate(pvalue1=pvalue) 
BW_GAM_mod_adj_GA$pvalue <-  format(round(BW_GAM_mod_adj_GA$pvalue, digits=2), nsmall=2)
BW_GAM_mod_adj_GA <- BW_GAM_mod_adj_GA%>% 
    mutate(pvalue1=case_when((pvalue1<0.0001)~ "<0.0001",
                             (pvalue1<0.001)~ "<0.001")) 
BW_GAM_mod_adj_GA$pvalue1 <- (ifelse(is.na(BW_GAM_mod_adj_GA$pvalue1),    BW_GAM_mod_adj_GA$pvalue,  BW_GAM_mod_adj_GA$pvalue1))
BW_GAM_mod_adj_GA <- BW_GAM_mod_adj_GA %>%
  select(-one_of('pvalue'))  
BW_GAM3_adj_GA <- BW_GAM_mod_adj_GA %>%
rename(pvalue=pvalue1) %>%
    mutate_if(is.numeric, ~ format(round(., digits = 2), nsmall = 2))

  BW_GAM3_adj_GA['Parameters'] <- c("intercept", 
                            "Disease during pregnancy (ref: none)"," ",
                           "Parity (ref: 1)", " ", "Sex (ref: male)",
                            "Morphology (ref: neither)", "", 
                            "Civil status (ref: married)",
                            "Lausanne (ref: yes)", ""
                            )
  BW_GAM3_adj_GA['category'] <- c(" ",
                          "influenza","syphilis",
                            "2", ">2", "female",
                            "Obese", "Thin",
                            "single or missing",
                           "no", "unsure")
    BW_GAM3_adj_GA<-BW_GAM3_adj_GA[, c('Parameters', "category", "beta", "lci","uci", "pvalue")]
    
pvalues_smooth_BW_adj_GA = summary(MBW.Z.GA2)
pvalues_smooth_BW_adj_GA <- pvalues_smooth_BW_adj_GA$s.table[,4]
pvalues_smooth_BW_adj_GA <- format(pvalues_smooth_BW_adj_GA, scientific=FALSE)

pvalues_smooth_BW_adj_GA_main_df <- data.frame(pvalues_smooth_BW_adj_GA) %>%
  dplyr::mutate(Parameters=c("Time", "Seasonality", "Maternal height (cm)", "Maternal years (age)")) %>%
  rename(c(pvalue=(pvalues_smooth_BW_adj_GA)))%>%
  mutate(pvalue=as.numeric(pvalue)) %>%
  mutate(pvalue1=pvalue) %>%
  add_column(beta="")%>%
  add_column(category="")%>%
  add_column(lci="") %>%
  add_column(uci="")

pvalues_smooth_BW_adj_GA_main_df$pvalue <-  format(round(pvalues_smooth_BW_adj_GA_main_df$pvalue, digits=2), nsmall=2)

pvalues_smooth_BW_adj_GA_main_df <- pvalues_smooth_BW_adj_GA_main_df %>%
  mutate(pvalue1=case_when((pvalue1<0.00001) ~ "<0.0001",
                           (pvalue1<0.001)~ "<0.001",
                           (pvalue1<0.01)~ "<0.01")) 

pvalues_smooth_BW_adj_GA_main_df$pvalue1 <- (ifelse(is.na(pvalues_smooth_BW_adj_GA_main_df$pvalue1), 
                                             pvalues_smooth_BW_adj_GA_main_df$pvalue, pvalues_smooth_BW_adj_GA_main_df$pvalue1))
pvalues_smooth_BW_adj_GA_main_df <- pvalues_smooth_BW_adj_GA_main_df%>%
  select(-one_of('pvalue')) %>%
  rename("pvalue"="pvalue1") 

rowinfo = c("Smooth variables", "", "", "", "", "")
ow_adj_BW_GA <- rbind(rowinfo,pvalues_smooth_BW_adj_GA_main_df)

BW_adj_GA <- rbind(BW_GAM3_adj_GA, ow_adj_BW_GA)

   MBW.Z.GA_gt <- BW_adj_GA %>%
      gt()%>%
      tab_spanner(label = "95% CI", columns = c(lci, uci),
                  ) %>%
     tab_style(
    style = list(
      cell_fill(color = "#e0e0e0")
    ),
    locations = cells_body(rows=c(2,3,6,9,13,15))) %>%
      tab_style(
    style = list(
      cell_fill(color = "#F2F0F0")),
    locations = cells_body(rows=c(4,5,7,8,10,11,14,16))) %>%
     tab_source_note(source_note = "n=7'533")  %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(rows = 12)) %>%
  tab_options(table.width = pct(25), data_row.padding = px(2),   table.font.size = 12,
              column_labels.font.weight = "bold",  row_group.font.weight  = "bold",
              heading.title.font.size = 14, heading.subtitle.font.size = 14
  )
    MBW.Z.GA_gt
    #pct 25
BW_estimates_flu_adj_GA_gt <- MBW.Z.GA_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP", "M_BW.Z.GA_estimates_flu_or_syph_adj_GA_gt.html"))
BW_estimates_flu_adj_GA_gt <- MBW.Z.GA_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP", "M_BW.Z.GA_estimates_flu_or_syph_adj_GA_gt.docx"))
```


#### Differentiating between live and stillbirths
##### M_BW.Z.score.GA_stillbirth
```{r}
M_BW_Z_GA_sb<- glm(Birthweight_Z_GA_LMP ~sex+ flu_or_syph*stillbirth, data=laus4, subset=(GA_weeks>22))
nobs(M_BW_Z_GA_sb)
summary(M_BW_Z_GA_sb) #n 8050
vif(M_BW_Z_GA_sb)
```

###### Quick summary table
```{r}
BW_GAM_flu_or_syph_mod <- round((estimates_cont(M_BW_Z_GA_sb)[c(1,3:7),]), 2)
BW_GAMmod <- BW_GAM_flu_or_syph_mod %>%
mutate(pvalue=as.numeric(pvalue)) %>%
mutate(pvalue1=pvalue) 
BW_GAMmod$pvalue <-  format(round(BW_GAMmod$pvalue, digits=2), nsmall=2)
BW_GAMmod <- BW_GAMmod%>% 
    mutate(pvalue1=case_when((pvalue1<0.0001)~ "<0.0001",
                             (pvalue1<0.001)~ "<0.001")) 
BW_GAMmod$pvalue1 <- (ifelse(is.na(BW_GAMmod$pvalue1),    BW_GAMmod$pvalue,  BW_GAMmod$pvalue1))
BW_GAMmod <- BW_GAMmod %>%
  select(-one_of('pvalue'))  
BW_GAM3 <- BW_GAMmod %>%
rename(pvalue=pvalue1) %>%
    mutate_if(is.numeric, ~ format(round(., digits = 2), nsmall = 2))

  BW_GAM3['Parameters'] <- c("intercept", "Disease during pregnancy (ref: none)","",
                             "Stillbirth (ref no)", "Influenza: yes * stillbirth (ref no)","Syphilis: yes * stillbirth (ref no)")
  BW_GAM3['category'] <- c("","influenza", "syphilis","yes", "stillbirth yes","stillbirth yes")
  BW_GAM3<-BW_GAM3[, c('Parameters', "category", "beta", "lci","uci", "pvalue")]
       
   BW_SB_GAM_gt <- BW_GAM3 %>%
      gt()%>%
      tab_spanner(label = "95% CI", columns = c(lci, uci))%>%
       tab_options(table.width = pct(25), data_row.padding = px(1)) %>%
     tab_source_note(source_note = "n=8'050")
    BW_SB_GAM_gt
    
M_BW.s_estimates_flu_gt <- BW_SB_GAM_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP/BW_HC_GA_interaction_stillbirth", "M_BW_Z_GA_sb_estimates_gt.html"))
M_BW.s_estimates_flu_gt <- BW_SB_GAM_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP/BW_HC_GA_interaction_stillbirth", "M_BW_Z_GA_sb_estimates_gt.docx"))
```


## Head circumference
#### Adjusting for gestational age
```{r}
MHC.Z.GA2 <- gam(head_circ_perturb_Z_GA_LMP ~ s(birthdate_num)+ s(day366, bs="cc") + s(height)+  s(age_mother)+flu_or_syph +parity_cat2+sex+morphology +civil_status_cat2+Lausanne, data=laus4_livebirths, subset = (GA_weeks>=22))
  summary(MHC.Z.GA2)
  # set.seed(1234)
  # gam.check(MHC.Z.GA, k.rep=2000)
gam.check(MHC.Z.GA2) 
concurvity(MHC.Z.GA2)
 BIC(MHC.Z.GA2)
 nobs(MHC.Z.GA2)
```

##### Summary table
```{r}
HC_GAM_mod_adj_GA <- round(head(estimates_GAM_cont(MHC.Z.GA2),11), 2)
HC_GAMmod_adj_GA <- HC_GAM_mod_adj_GA %>%
mutate(pvalue=as.numeric(pvalue)) %>%
mutate(pvalue1=pvalue) 
HC_GAMmod_adj_GA$pvalue <-  format(round(HC_GAMmod_adj_GA$pvalue, digits=2), nsmall=2)
HC_GAMmod_adj_GA <- HC_GAMmod_adj_GA%>% 
    mutate(pvalue1=case_when((pvalue1<0.0001)~ "<0.0001",
                             (pvalue1<0.001)~ "<0.001")) 
HC_GAMmod_adj_GA$pvalue1 <- (ifelse(is.na(HC_GAMmod_adj_GA$pvalue1),    HC_GAMmod_adj_GA$pvalue,  HC_GAMmod_adj_GA$pvalue1))
HC_GAMmod_adj_GA <- HC_GAMmod_adj_GA %>%
  select(-one_of('pvalue'))  
HC_GAM3_adj_GA <- HC_GAMmod_adj_GA %>%
rename(pvalue=pvalue1) %>%
    mutate_if(is.numeric, ~ format(round(., digits = 2), nsmall = 2))

  HC_GAM3_adj_GA['Parameters'] <- c("intercept", 
                            "Disease during pregnancy (ref: none)"," ",
                            "Parity (ref: 1)", " ", "Sex (ref: male)",
                            "Morphology (ref: neither)", "", 
                            "Civil status (ref: married)",
                            "Lausanne (ref: yes)", ""
                            )
  HC_GAM3_adj_GA['category'] <- c(" ",
                          "influenza","syphilis",
                            "2", ">2", "female", 
                            "Obese", "Thin",
                            "single or missing",
                           "no", "unsure")
    HC_GAM3_adj_GA<-HC_GAM3_adj_GA[, c('Parameters', "category", "beta", "lci","uci", "pvalue")]
    
pvalues_smooth_HC_adj_GA = summary(MHC.Z.GA2)
pvalues_smooth_HC_adj_GA <- pvalues_smooth_HC_adj_GA$s.table[,4]
pvalues_smooth_HC_adj_GA <- format(pvalues_smooth_HC_adj_GA, scientific=FALSE)

pvalues_smooth_HC_adj_GA_main_df <- data.frame(pvalues_smooth_HC_adj_GA) %>%
  dplyr::mutate(Parameters=c("Time", "Seasonality", "Maternal height (cm)", "Maternal years (age)")) %>%
  rename(c(pvalue=(pvalues_smooth_HC_adj_GA)))%>%
  mutate(pvalue=as.numeric(pvalue)) %>%
  mutate(pvalue1=pvalue) %>%
  add_column(beta="")%>%
  add_column(category="")%>%
  add_column(lci="") %>%
  add_column(uci="")

pvalues_smooth_HC_adj_GA_main_df$pvalue <-  format(round(pvalues_smooth_HC_adj_GA_main_df$pvalue, digits=2), nsmall=2)

pvalues_smooth_HC_adj_GA_main_df <- pvalues_smooth_HC_adj_GA_main_df %>%
  mutate(pvalue1=case_when((pvalue1<0.00001) ~ "<0.0001",
                           (pvalue1<0.001)~ "<0.001",
                           (pvalue1<0.01)~ "<0.01")) 

pvalues_smooth_HC_adj_GA_main_df$pvalue1 <- (ifelse(is.na(pvalues_smooth_HC_adj_GA_main_df$pvalue1), 
                                             pvalues_smooth_HC_adj_GA_main_df$pvalue, pvalues_smooth_HC_adj_GA_main_df$pvalue1))
pvalues_smooth_HC_adj_GA_main_df <- pvalues_smooth_HC_adj_GA_main_df%>%
  select(-one_of('pvalue')) %>%
  rename("pvalue"="pvalue1") 

rowinfo = c("Smooth variables", "", "", "", "", "")
ow_adj_GA <- rbind(rowinfo,pvalues_smooth_HC_adj_GA_main_df)

HC_adj_GA <- rbind(HC_GAM3_adj_GA, ow_adj_GA)

   MHC.Z.GA_gt <- HC_adj_GA %>%
      gt()%>%
      tab_spanner(label = "95% CI", columns = c(lci, uci),
                  ) %>%
     tab_style(
    style = list(
      cell_fill(color = "#e0e0e0")
    ),
    locations = cells_body(rows=c(2,3,6,9,13,15))) %>%
      tab_style(
    style = list(
      cell_fill(color = "#F2F0F0")),
    locations = cells_body(rows=c(4,5,7,8,10,11,14,16))) %>%
     tab_source_note(source_note = "n=7'504")  %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(rows = 12)) %>%
  tab_options(table.width = pct(18), data_row.padding = px(2),   table.font.size = 12,
              column_labels.font.weight = "bold",  row_group.font.weight  = "bold",
              heading.title.font.size = 14, heading.subtitle.font.size = 14
  )
    MHC.Z.GA_gt
    
HC_estimates_flu_adj_GA_gt <- MHC.Z.GA_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP", "M_HC.Z.GA_estimates_flu_or_syph_adj_GA_gt.html"))
HC_estimates_flu_adj_GA_gt <- MHC.Z.GA_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP", "M_HC.Z.GA_estimates_flu_or_syph_adj_GA_gt.docx"))
```


##### M_HC_Z.score_GA_stillbirth
```{r}
M_HC_Z_GA_sb <- glm(head_circ_perturb_Z_GA_LMP ~sex+ flu_or_syph*stillbirth, data=laus4, subset = (GA_weeks>=22))
summary(M_HC_Z_GA_sb)
nobs(M_HC_Z_GA_sb)#n 7,924
vif(M_HC_Z_GA_sb)
```

###### Quick summary table
```{r}
HC_GAM_syph_flu_mod <- round((estimates_cont(M_HC_Z_GA_sb)[c(1,3:7),]), 2)
HC_GAMmod <- HC_GAM_syph_flu_mod %>%
mutate(pvalue=as.numeric(pvalue)) %>%
mutate(pvalue1=pvalue) 
HC_GAMmod$pvalue <-  format(round(HC_GAMmod$pvalue, digits=2), nsmall=2)
HC_GAMmod <- HC_GAMmod%>% 
    mutate(pvalue1=case_when((pvalue1<0.0001)~ "<0.0001",
                             (pvalue1<0.001)~ "<0.001")) 
HC_GAMmod$pvalue1 <- (ifelse(is.na(HC_GAMmod$pvalue1),    HC_GAMmod$pvalue,  HC_GAMmod$pvalue1))
HC_GAMmod <- HC_GAMmod %>%
  select(-one_of('pvalue'))  
HC_GAM3 <- HC_GAMmod %>%
rename(pvalue=pvalue1) %>%
    mutate_if(is.numeric, ~ format(round(., digits = 2), nsmall = 2))

  HC_GAM3['Parameters'] <- c("intercept", 
                             "Disease during pregnancy (ref: none)", 
                             " ",
                             "Stillbirth (ref no)", "Influenza: yes * stillbirth (ref no)", "Syphilis: yes * stillbirth (ref no)")
  HC_GAM3['category'] <- c("","influenza", "syphilis", "yes","stillbirth yes","stillbirth yes")
  HC_GAM3<-HC_GAM3[, c('Parameters', "category", "beta", "lci","uci", "pvalue")]
    
   HC_SB_GAM_flu_syph_gt <- HC_GAM3 %>%
      gt()%>%
      tab_spanner(label = "95% CI", columns = c(lci, uci))%>%
       tab_options(table.width = pct(25), data_row.padding = px(1)) %>%
     tab_source_note(source_note = "n=7'924")
    HC_SB_GAM_flu_syph_gt
    # pct 20
HC_estimates_flu_gt <- HC_SB_GAM_flu_syph_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP/BW_HC_GA_interaction_stillbirth", "M_HC_Z_GA_sb_gt_estimates.html"))
HC_estimates_flu_gt <- HC_SB_GAM_flu_syph_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP/BW_HC_GA_interaction_stillbirth", "M_HC_Z_GA_sb_gt_estimates.docx"))
```

# Gestational age
```{r}
M_GA<- gam(GA_weeks ~ s(birthdate_num)+ s(day366, bs="cc") + s(height)+  s(age_mother) +flu_or_syph+parity_cat2+sex+morphology +civil_status_cat2+Lausanne, data=laus4_livebirths, subset=(GA_weeks>=22))
  summary(M_GA)
  gam.check(M_GA)
  concurvity(M_GA)
nobs(M_GA) #n=7538
```

## Summary table
```{r}
GA_GAM_mod <- round(head(estimates_GAM_cont(M_GA),11), 2)
GA_GAMmod <- GA_GAM_mod %>%
mutate(pvalue=as.numeric(pvalue)) %>%
mutate(pvalue1=pvalue) 
GA_GAMmod$pvalue <-  format(round(GA_GAMmod$pvalue, digits=2), nsmall=2)
GA_GAMmod <- GA_GAMmod%>% 
    mutate(pvalue1=case_when((pvalue1<0.0001)~ "<0.0001",
                             (pvalue1<0.001)~ "<0.001")) 
GA_GAMmod$pvalue1 <- (ifelse(is.na(GA_GAMmod$pvalue1),    GA_GAMmod$pvalue,  GA_GAMmod$pvalue1))
GA_GAMmod <- GA_GAMmod %>%
  select(-one_of('pvalue'))  
GA_GAM3 <- GA_GAMmod %>%
rename(pvalue=pvalue1) %>%
    mutate_if(is.numeric, ~ format(round(., digits = 2), nsmall = 2))

  GA_GAM3['Parameters'] <- c("intercept", 
                            "Disease during pregnancy (ref: no)"," ", 
                            "Parity (ref: 1)", " ", "Sex (ref: male)",
                            "Morphology (ref: neither)", "", 
                            "Civil status (ref: married)",
                            "Lausanne (ref: yes)", ""
                            )
  GA_GAM3['category'] <- c(" ",
                          "influenza","syphilis",  
                            "2", ">2", "female", 
                            "Obese", "Thin",
                            "single or missing",
                           "no", "unsure")
    GA_GAM3<-GA_GAM3[, c('Parameters', "category", "beta", "lci","uci", "pvalue")]
    
pvalues_smooth_GA_main = summary(M_GA)
pvalues_smooth_GA_main <- pvalues_smooth_GA_main$s.table[,4]
pvalues_smooth_GA_main <- format(pvalues_smooth_GA_main, scientific=FALSE)

pvalues_smooth_GA_main_df <- data.frame(pvalues_smooth_GA_main) %>%
  dplyr::mutate(Parameters=c("Time", "Seasonality", "Maternal height (cm)", "Maternal years (age)")) %>%
  rename(c(pvalue=(pvalues_smooth_GA_main)))%>%
  mutate(pvalue=as.numeric(pvalue)) %>%
  mutate(pvalue1=pvalue) %>%
  add_column(beta="")%>%
  add_column(category="")%>%
  add_column(lci="") %>%
  add_column(uci="")

pvalues_smooth_GA_main_df$pvalue <-  format(round(pvalues_smooth_GA_main_df$pvalue, digits=2), nsmall=2)

pvalues_smooth_GA_main_df <- pvalues_smooth_GA_main_df %>%
  mutate(pvalue1=case_when((pvalue1<0.00001) ~ "<0.0001",
                           (pvalue1<0.001)~ "<0.001",
                           (pvalue1<0.01)~ "<0.01")) 

pvalues_smooth_GA_main_df$pvalue1 <- (ifelse(is.na(pvalues_smooth_GA_main_df$pvalue1), 
                                             pvalues_smooth_GA_main_df$pvalue, pvalues_smooth_GA_main_df$pvalue1))
pvalues_smooth_GA_main_df <- pvalues_smooth_GA_main_df%>%
  select(-one_of('pvalue')) %>%
  rename("pvalue"="pvalue1") 

rowinfo = c("Smooth variables", "", "", "", "", "")
ow <- rbind(rowinfo,pvalues_smooth_GA_main_df)

GA <- rbind(GA_GAM3, ow)

   M_GA_gt <- GA %>%
      gt()%>%
      tab_spanner(label = "95% CI", columns = c(lci, uci),
                  ) %>%
     tab_style(
    style = list(
      cell_fill(color = "#e0e0e0")
    ),
    locations = cells_body(rows=c(2,3,6,9,13,15))) %>%
      tab_style(
    style = list(
      cell_fill(color = "#F2F0F0")),
    locations = cells_body(rows=c(4,5,7,8,10,11,14,16))) %>%
     tab_source_note(source_note = "n=7'538")  %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(rows = 12)) %>%
  tab_options(table.width = pct(25), data_row.padding = px(2),   table.font.size = 12,
              column_labels.font.weight = "bold",  row_group.font.weight  = "bold",
              heading.title.font.size = 14, heading.subtitle.font.size = 14
  )
    M_GA_gt
    
GA_estimates_flu_gt <- M_GA_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP", "M_GA_estimates_flu_or_syph_gt.html"))
GA_estimates_flu_gt <- M_GA_gt%>%
  gtsave(here("outputs/tables/suppl_data/models_with_GA_based_on_LMP", "M_GA_estimates_flu_or_syph_gt.docx"))
```

# Anthropometrics sex - Z-score and infections
## Overall
# stillbirth and GA
```{r}
sb_ga <- laus4 %>%
  group_by(stillbirth, flu_or_syph) %>%
  filter(!is.na(hc.perturbed) & !is.na(GA_weeks))%>%
  summarise(mean(birthweight), sd(birthweight), 
            mean(hc.perturbed), sd(hc.perturbed),
            mean(GA_weeks), sd(hc.perturbed))
sb_ga

sb_ga1 <- laus4 %>%
  group_by(stillbirth, flu_or_syph) %>%
  filter(!is.na(hc.perturbed)& !is.na(GA_weeks))%>%
  summarise(mean(birthweight), mean(hc.perturbed), mean(GA_weeks))
sb_ga1
```

## Plot
```{r}
BW_GA_SB_inf <- sb_ga %>%
ggplot(aes(x = `mean(GA_weeks)`, y = `mean(birthweight)`, 
           color = stillbirth, shape=stillbirth)) +
   geom_point(size=4) +
  labs(x = "Gestational age (weeks)", y = "Birth weight (g)") +
  scale_color_manual(values = c("livebirth" = "#009E73", "stillbirth" = "#D55E00")) +
  scale_shape_manual(values = c("livebirth" = 19, "stillbirth" = 17)) +
 scale_y_continuous(breaks=seq(1700,3300, by=200))+
  scale_x_continuous(breaks=seq(33,40, by=1))+
 coord_cartesian(xlim=c(33,40), ylim = c(1700,3300))+
theme_minimal() +
    geom_line(data = subset(sb_ga, flu_or_syph == "no"), 
            aes(group = 1), 
            linetype = "solid", colour="#56B4E9", linewidth=0.8) +
   geom_line(data = subset(sb_ga, flu_or_syph == "influenza"), 
            aes(group = 1), 
            linetype = "solid", colour="#CC79A7", linewidth=0.8) +
   geom_line(data = subset(sb_ga, flu_or_syph == "syphilis"), 
            aes(group = 1), 
            linetype = "solid", colour="#E69F00", linewidth=0.8)+
    annotate("text",colour="#E69F00",x=34.8,y=2180,
             label="syphilis")+
    annotate("text",colour="#CC79A7",x=37,y=2670,
             label="influenza")+
    annotate("text",colour="#56B4E9",x=38.3,y=3100,
             label="no disease")+
  theme(legend.position="none")+
  custom_theme()+
  NULL
BW_GA_SB_inf


HC_GA_SB_inf <- sb_ga %>%
ggplot(aes(x = `mean(GA_weeks)`, y = `mean(hc.perturbed)`, 
           color = stillbirth, shape=stillbirth)) +
   geom_point(size=4) +
  scale_y_continuous(breaks=seq(27,35,by=1)) +
  scale_x_continuous(breaks=seq(33,40,by=1)) +
  coord_cartesian(xlim=c(33,40),ylim=c(27,35))+
  labs(x = "Gestational age (weeks)", y = "Head circumference (cm)") +
  scale_color_manual(values = c("livebirth" = "#009E73", "stillbirth" = "#D55E00")) +
  scale_shape_manual(values=c("livebirth"=19, "stillbirth"=17))+
   theme_minimal() +
    geom_line(data = subset(sb_ga, flu_or_syph == "no"), 
            aes(group = 1), 
            linetype = "solid", colour="#56B4E9", linewidth=0.8) +
   geom_line(data = subset(sb_ga, flu_or_syph == "influenza"), 
            aes(group = 1), 
            linetype = "solid", colour="#CC79A7", linewidth=0.8) +
   geom_line(data = subset(sb_ga, flu_or_syph == "syphilis"), 
            aes(group = 1), 
            linetype = "solid", colour="#E69F00", linewidth=0.8) +
    annotate("text",colour="#E69F00",x=35.2,y=30,
             label="syphilis")+
    annotate("text",colour="#CC79A7",x=37,y=32.2,
             label="influenza")+
    annotate("text",colour="#56B4E9",x=38.4,y=34.2,
             label="no disease")+
  custom_theme()+
  NULL
HC_GA_SB_inf


BW_HC_GA_SB_inf <- egg::ggarrange(BW_GA_SB_inf, 
          HC_GA_SB_inf,
          labels = c("A", "B"),
          ncol = 2, nrow = 1) 
# ggsave("output/graphs/brain_sparing/infections_BW_HC_stillbirth_GA.pdf", BW_HC_GA_SB_inf, width = 16, height = 7, units = "in", device = "pdf")
```
